---
title: Fixed size clustering
output:
    rmarkdown::html_vignette:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Introduction to fixed size clustering}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
Sys.setenv(LANGUAGE = "en")
library("minair")
```

# Fixed size clustering

In general clustering algorithms minimize/maximize some metric in a cluster
ignoring its size.
In rare circumstance the size of the clusters has to be similar or fixed.
Here we demonstrate modified cluster algorithms to support fixed sized clusters.

```{r prepare-data}
mtx <- scale(iris[, -5])
dst <- dist(mtx, method = "euclidean")
```

# clustering

Here we run three different restricted clustering alogrithms against the
classical partitioning algorithm `kmeans`.

```{r clustering}
set.seed(123)
cluster_size <- 40

hcf <- fixed_size_clustering_hclust(dst, size = cluster_size)
nnf<- fixed_size_clustering_nearest_neighbor(dst, size = cluster_size)
kmf <- fixed_size_clustering_kmeans(mtx, size = cluster_size)
kmd <- kmeans(mtx, centers = ceiling(nrow(mtx) / cluster_size))
```

## Plot dendrogram

```{r plot-dendrogram, message = FALSE}
library("dendextend")

dend <- as.dendrogram(hclust(dst, method = "ward.D2"))
cluster_ids <- cbind(
    "hclust_fixed" = hcf,
    "nrstnb_fixed" = nnf,
    "kmeans_fixed" = kmf,
    "kmeans_default" = kmd$cluster
)

par(mar = c(8, 4, 4, 2) + .1)
plot(dend)
colored_bars(cluster_ids, dend)
```

## Clustering performance

```{r wcss}
wss <- sapply(
    list(hcf, nnf, kmf, kmd$cluster), within_cluster_sum_of_squares, x = mtx
)
tss <- total_sum_of_squares(mtx)
colnames(wss) <- colnames(cluster_ids)

col <- seq_len(ncol(wss))
barplot(
    t(wss), col = col, beside = TRUE,
    main = "Within cluster sum of squares"
)
legend("topright", legend = colnames(wss), col = col, pch = 15)
```

```{r b-tss}
barplot(
    1 - colSums(wss) / tss, col = col, horiz = TRUE,
    main = "Between/Total sum of squares"
)
legend(
    "bottomright",
    legend = colnames(wss),
    col = col, pch = 15,
    bty = "n"
)
```

# Session Information

```{r sessionInfo}
sessionInfo()
```
